---
title: "Homework 1"
author: Gauri Bhatkhande
output: github_document

---

This is my solution to HW1 

*Loaded the packages that I need for this project.* 
```{r}
library(tidyverse)
```
```{r}
library(palmerpenguins)
```


## Problem 1

*Create a dataframe comprised of the specified elements:* 

```{r}
prob1_df = 
tibble(
    samp = rnorm(10),
    samp_gt_0 = samp > 0, 
    char_vec = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    factor_vec = factor(c("low", "low", "low", "mod", "mod", "mod", "mod", "high", "high", "high" ))
)
```

*Take the mean of each variable in my data frame.* 

```{r}
mean(pull(prob1_df, samp))
mean(pull(prob1_df, samp_gt_0))
mean(pull(prob1_df, char_vec))
mean(pull(prob1_df, factor_vec))
```
I can take the means of numbers and logical but not character or factor. Logical gets understood by R as 1 or 0 and hence R at the backend does that calculation for us and gives the mean 

```{r results=FALSE}
as.numeric(pull(prob1_df, samp_gt_0))
as.numeric(pull(prob1_df, char_vec))
as.numeric(pull(prob1_df, factor_vec))
```
Can convert the logical and factor vectors to numerics and then take means but could not convert the character vector to numeric and hence could not take the mean 

```{r}
as.numeric(pull(prob1_df, samp_gt_0)) * pull(prob1_df, samp)
```

```{r}
as.factor(pull(prob1_df, samp_gt_0)) * pull(prob1_df, samp)
```

```{r}
as.numeric(as.factor(pull(prob1_df, samp_gt_0))) * pull(prob1_df, samp)
```

## Problem 2 

*I loaded the penguins dataset:* 
```{r}
data("penguins", package = "palmerpenguins")

```
*Description of the dataset:*

```{r}
nrow(penguins)
ncol(penguins)
```

There are **344 observations** and **8 variables** in the dataset *penguins*

*I want to explore which are the 8 variable names:* 
```{r}
variable.names(penguins)
```
The names of the variables in the dataset are *species, island, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g, sex, year*

*To find the categories of species and islands and the number of penguins pertaining to each species and island:* 
```{r}
levels(pull(penguins, species))
levels(pull(penguins, island))
```
```{r}
count(penguins, species)
count(penguins, island)
count(penguins, sex)
```

The three penguin species are *Adelie (n = 152), Chinstrap (n = 68), Gentoo (n=124)*
The islands are named *Biscoe (n = 168), Dream (n = 124) and Torgersen (n = 52)*
*165* penguins are *females*, *168* are *males* and *11* are *unknown*.

*To find the mean size measurements:*
 
```{r}

mean (pull(penguins, bill_length_mm),na.rm = TRUE)
mean (pull(penguins, bill_depth_mm),na.rm = TRUE)
mean (pull(penguins, flipper_length_mm),na.rm = TRUE)
mean (pull(penguins, body_mass_g),na.rm = TRUE)

```
The mean *bill length* for the penguins is *43.92 mm.* 
The mean *bill depth* for the penguins is *17.15 mm.* 
The mean *flipper length* for the penguins is *200.91 mm.* 
The mean *body mass* for the penguins is *4201.75 grams.* 

I wanted to know the mean size measurements which I believe are important here and stratified them by the species group.
*The mean size measurements stratified by the species are calculated below:*
```{r}
aggregate(penguins[,3:6],list(pull(penguins,species)),mean,na.rm=TRUE)
```

I wanted to know whether the mean sizes for a species differed by the islands. 
*Therefore I stratified that relationship with islands.*
```{r}
aggregate(penguins[,3:6],list(pull(penguins,island), pull(penguins,species)),mean,na.rm=TRUE)

```


*I calculated the number of rows and columns in the dataset:*
```{r}
nrow(penguins)

```

There are **344 rows** in the *penguins* dataset

```{r}
ncol(penguins)
```
There are **8 columns** in the *penguins* dataset 

```{r}
mean (as.numeric(pull(penguins, flipper_length_mm)),na.rm = TRUE)
```

**The mean flipper length is 200.9152 mm.** 

*I made the following scatterplot:*
```{r}
y = pull(penguins, flipper_length_mm)
x = pull(penguins, bill_length_mm)
ggplot(penguins, aes(x = x, y = y, color = species)) + geom_point() + labs(x="Bill length", y = "Flipper length") + ggtitle("Scatterplot of flipper length vs bill length of penguins")
```

*Received the following Warning message:Removed 2 rows containing missing values (geom_point). This could be because two rows had NA for flipper length*

```{r}
ggsave("scatter_plot1.pdf" )
```

